Ant Transition Matrix

### 2020-10-20 Cholla_Analysis.Rmd
initial commit of Ali's `Shared Code' folder contents.

@@

```

## Vital Rate Plots
#### Survival Rate

Three models were considered here. First A simple survival vs log(volume) model. This had an AIC value of 2380.825.  The second included a second explanatory variable of the ant species (as an independent variable). This had an AIC vaue of 2474.975/ The third included the ant species variable as an interaction term with the volume. This had an AIC value of 2378.601, which is a difference greater than 2, and therefore is a satistically supported improvement. 
```{r}
surv_model <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
```

@@

This includes an ant interaction term
```{r}
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)

#anova(surv_model_ant_interact, test = "Chisq")
#summary(surv_model_ant_interact)
```

```{r}
ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0.1) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

#### Flowering Model
```{r}
flo_model <- glm(Goodbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)

cactus$Flow <- cactus$Goodbuds_t > 0 #blank column 
for(i in 1:8189){
  if(cactus$TotFlowerbuds_t[i] == NA){
    cactus$Flow[i] <- sum(cactus$ABFlowerbuds_t[i]+cactus$Goodbuds_t[i])
  }
  else if(cactus$TotFlowerbuds_t[i] == 0){
    cactus$Flow[i] <- 0
  }
  else if(cactus$TotFlowerbuds_t[i] >0){
    cactus$Flow[i] <- 1
  }
}

@@

#### Growth Model
```{r}
gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                  data=cactus)

ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=as.factor(bin1))) +  geom_point(alpha = 0) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) 
```
## Ant Transition Model
```{r}
ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

## Multinomial Models

@@

>>>>

## Ant Transition Model
```{r}
ant_trans_model <- glm(as.factor(bin2) ~ 0 + as.factor(bin1),family = binomial,data=cactus)
ggplot(cactus, aes(y=as.factor(bin2), x=as.factor(bin1), col=as.factor(bin1))) +  geom_point(alpha = 0.9) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) 
```

@@

```
```{r}
survival_plot <- cactus %>% 
  mutate(size_bin = cut_interval(log(volume_t),10)) %>% 
  group_by(size_bin, cactus$bin1) %>% 
  summarise(mean_size = mean(log(volume_t),na.rm=T),
    surv = mean(Survival_t1,na.rm=T))

plot(log(cactus$volume_t),cactus$Survival_t1)
points(survival_plot$mean_size,survival_plot$surv,pch=16,cex=1, col=1:length(cactus$bin1))
legend(12,1,unique(cactus$bin1),col=1:length(cactus$bin1),pch=1)
```

## Vital Rate Binomial Models
#### Survival Rate
Three models were considered here. First A simple survival vs log(volume) model. This had an AIC value of 2380.825.  The second included a second explanatory variable of the ant species (as an independent variable). This had an AIC vaue of 2474.975/ The third included the ant species variable as an interaction term with the volume. This had an AIC value of 2378.601, which is a difference greater than 2, and therefore is a satistically supported improvement. 
```{r}
surv_model <- glm(Survival_t1 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
surv_model_ant <- glm(Survival_t1 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
```

@@

This includes an ant interaction term
```{r}
surv_model_ant_interact <- glm(Survival_t1 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)

anova(surv_model_ant_interact, test = "Chisq")
summary(surv_model_ant_interact)
```

```{r}
ggplot(cactus, aes(y=Survival_t1, x=log(volume_t), col=factor(bin1, label = c("Liom","Crem","For","Vacant","Other")))) +  geom_point(alpha = 0.1) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', method.args = list(family = "binomial"), se = FALSE) +
  labs(title = "Survival Rate",colour = "Ant Species") + 
  xlab("Volume at t") + 
  ylab("Survival Rate")
```

#### Flowering Model
The three models considered here model the flowering rates. The first considers only the flowering rates vs the log(volume). The AIC value for this model is 246.7699. The second includes the ant species as a second explanatory variable. The AIC value for this is 250.4667. The third model includes the ant species as an interacting variable. The AIC value for this is 254.5773. Therefore there is no statistical evidence that the models including the ant species, either as an independent or dependent variable, are more relevant. 
```{r}
flo_model <- glm(TotFlowerbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
flo_model_ant <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) + as.factor(bin1),
                  family="binomial",
                  data=cactus)
flo_model_ant_interact <- glm(TotFlowerbuds_t > 0 ~ 0 + log(volume_t) * as.factor(bin1),
                  family="binomial",
                  data=cactus)
```

```{r}
flo_model <- glm(Goodbuds_t > 0 ~ log(volume_t),
                  family="binomial",
                  data=cactus)
summary(flo_model)
```

```{r}
cactus$Flow <- cactus$Goodbuds_t > 0 #blank column 
for(i in seq_len(nrow(cactus))){
  if(na.omit(cactus$TotFlowerbuds_t[i]) > 0){cactus$Flow[i] <- 1}
}

na.omit(cactus$TotFlowerbuds_t[100]) > 0

@@

#### Growth Model
Three models were considered for the growth rates vs the log(volume) of the cacti. The first is just the log(volue) of the second year vs the log(volume) of the first. The AIC value for this is 16126.65. The second includes the ant species as a second explanatory variable. The AIC value for this is 16039.89. The third model includes the ant species as an interacting variable. The AIC value for this is 16007.67. This means that the third model is statistically improved from the other two. 
```{r}
gro_model <- glm(log(volume_t1) ~ log(volume_t),
                  data=cactus)
gro_model_ant <- glm(log(volume_t1) ~ 0 + log(volume_t) + as.factor(bin1),
                  data=cactus)
gro_model_ant_interact <- glm(log(volume_t1) ~ 0 + log(volume_t) * as.factor(bin1),
                  data=cactus)
```

```{r}
#summary(gro_model_ant_interact)
ggplot(cactus, aes(y=log(volume_t1), x=log(volume_t), col=factor(bin1, label = c("Liom","Crem","For","Vacant","Other")))) +  geom_point(alpha = 0) +
  geom_smooth(data=cactus, method = "glm", formula='y~x', se = FALSE) +
  labs(title = "Growth Rate",colour = "Ant Species") + 
  xlab("Volume at t") + 
  ylab("Volume at t1")
```

```{r}
par(mfrow=c(2,2),mar=c(4,4,2,1))
xx=seq(0,8,by=.01)
hist(cactus$volume_t1[is.na(cactus$volume_t)],main='Size of Recruits',freq=FALSE)
   # lines(xx,dnorm(xx,mean(cactus$volume_t1[is.na(cactus$volume_t)]),sd(cactus$volume_t1[is.na(cactus$volume_t)])),col='red',lwd=3)
```

## Multinomial Models of Vital Rates

@@

##Vital Rate Functions
List of Parameters:
```{r}
params=data.frame(
    ##Survival Functions (by ant bin)
    surv1.int=NA,
    surv1.slope=NA,
    surv2.int=NA,
    surv2.slope=NA,
    surv3.int=NA,
    surv3.slope=NA,
    surv4.int=NA,
    surv4.slope=NA,
    surv5.int=NA,
    surv5.slope=NA,
    ##Growth Function
    flo.int=NA,
    flo.slope=NA,
    ##Flower Function
    gro1.int=NA,
    gro1.slope=NA,
    gro2.int=NA,
    gro2.slope=NA,
    gro3.int=NA,
    gro3.slope=NA,
    gro4.int=NA,
    gro4.slope=NA,
    gro5.int=NA,
    gro5.slope=NA
)
params$flo.int = coef(flo_model)[1]
params$flo.slope = coef(flo_model)[2]
```

Probability of Surviving:
```{r}
s1.x=function(x,params) {
    u=exp(params$surv1.int+params$surv1.slope*x)
    return(u/(1+u))
}
s2.x=function(x,params) {
    u=exp(params$surv2.int+params$surv2.slope*x)
    return(u/(1+u))
}
s3.x=function(x,params) {
    u=exp(params$surv3.int+params$surv3.slope*x)
    return(u/(1+u))
}
s4.x=function(x,params) {
    u=exp(params$surv4.int+params$surv4.slope*x)
    return(u/(1+u))
}
s5.x=function(x,params) {
    u=exp(params$surv5.int+params$surv5.slope*x)
    return(u/(1+u))
}
```

Probability of Flowering:
```{r}
f.x=function(x,params) {
    u=exp(params$flo.int+params$flo.slope*x)
    return(u/(1+u))
}
```

Probability of Growth:
```{r}
g1.x=function(x,params){
    u=exp(params$growth1.int+params$growth1.slope*x)
    return(u/(1+u))
}
```
